Social Housing in Toronto

### 2024-04-22 paper/paper.qmd
commit

To clean the data for better analysis processing, unnecessary columns such as how many bedrooms are in the social housing units and repetitive columns were removed. We use the function 'janitor' to achieve intended data cleaning.

@@

```{r}
#| label: applications
#| fig-cap: Data Overview
#| echo: false

@@

The above @figure1 is an overview of the data. It simulates the first 6 lines of all columns to give the readers an idea of what the data looks like. The complete data consists of 9 columns/variables and 20 lines/quarters of years from quarter 1 2019 to quarter 4 2023.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

```
The dot graph @figure2 illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023. The data exhibits fluctuating patterns in new or reactivated social housing applications across different quarters, with a notable surge observed in Q1 2020. Despite these fluctuations, the number of households housed each quarter consistently lags behind new entries, underscoring a substantial gap between demand and supply dynamics. This persistent shortfall indicates the ongoing challenge of meeting the housing needs of vulnerable populations within the community.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

```
The dot plot @figure3 comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time. From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period. However, seniors' applications appear to comprise a consistantly smaller proportion of the total active applications, as indicated by the consistently lower values for seniors' applications compared to total active applications. This disparity suggests that while seniors constitute a significant demographic group in need of social housing support, they may not represent the majority of applicants. Understanding this dynamic is crucial for policymakers and housing providers to tailor their strategies and resources effectively to address the diverse needs of applicants across different age groups.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

The dot plot @figure4 comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time. From the graph, we observe fluctuations in both the number of households with and without dependents applying for social housing throughout the observed period. Interestingly, the number of households with dependents appears to fluctuate more prominently compared to households without dependents, as indicated by the greater variability in the dashed line representing households with dependents.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

```
The dot plot @figure5 visualizes the relationship between inactive/cancelled applications and the total applications on the waiting list in Toronto's social housing sector from Q1 2019 through Q4 2023. The graph demonstrates the fluctuating nature of inactive or cancelled applications across various quarters, with notable spikes and declines that provide insights into the dynamics affecting the housing waiting list. For example, a substantial rise in cancellations or inactive statuses during specific periods, which are quarter 2 2019 and quarter 1 2021, could indicate shifts in applicant circumstances or dissatisfaction with the waiting times, which in turn impacts the overall efficiency of housing allocation. These fluctuations highlight the challenges in maintaining an active and accurate waiting list, emphasizing the need for robust management systems to handle the complexities of application statuses.

@@

\appendix

# Appendix {-}

>>>>

To clean the data for better analysis processing, unnecessary columns such as how many bedrooms are in the social housing units and the repetitive column 'total application' were removed. We use the function 'janitor' to achieve intended data cleaning.

@@

```{r}
#| label: data-overview
#| #| fig-cap: Data Overview
#| fig.pos: H
#| label: applications
#| fig-cap: Data Overview
#| echo: false

@@

The above @data-overview is an overview of the data. It simulates the first 6 lines of all columns to give the readers an idea of what the data looks like. The complete data consists of 9 columns/variables and 20 lines/quarters of years from quarter 1 2019 to quarter 4 2023.

@@

```{r}
#| label: new-and-housed
#| fig.pos: H
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

```
The dot plot @new-and-housed illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023. The data exhibits fluctuating patterns in new or reactivated social housing applications across different quarters, with a notable surge observed in Q1 2020. Despite these fluctuations, the number of households housed each quarter consistently lags behind new entries, underscoring a substantial gap between demand and supply dynamics. This persistent shortfall indicates the ongoing challenge of meeting the housing needs of vulnerable populations within the community.

@@

```{r}
#| label: total-and-seniors
#| fig.pos: H
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

```
The dot plot @total-and-seniors comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time. From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period. However, seniors' applications appear to comprise a consistantly smaller proportion of the total active applications, as indicated by the consistently lower values for seniors' applications compared to total active applications. This disparity suggests that while seniors constitute a significant demographic group in need of social housing support, they may not represent the majority of applicants. Understanding this dynamic is crucial for policymakers and housing providers to tailor their strategies and resources effectively to address the diverse needs of applicants across different age groups.

@@

```{r}
#| label: dependents-or-not
#| fig.pos: H
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

The dot plot @dependents-or-not comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time. From the graph, we observe fluctuations in both the number of households with and without dependents applying for social housing throughout the observed period. Interestingly, the number of households with dependents appears to fluctuate more prominently compared to households without dependents, as indicated by the greater variability in the dashed line representing households with dependents.

@@

```{r}
#| label: cancelled-vs-total
#| fig.pos: H
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

```
The dot plot @cancelled-vs-total visualizes the relationship between inactive/cancelled applications and the total applications on the waiting list in Toronto's social housing sector from Q1 2019 through Q4 2023. The graph demonstrates the fluctuating nature of inactive or cancelled applications across various quarters, with notable spikes and declines that provide insights into the dynamics affecting the housing waiting list. For example, a substantial rise in cancellations or inactive statuses during specific periods, which are quarter 2 2019 and quarter 1 2021, could indicate shifts in applicant circumstances or dissatisfaction with the waiting times, which in turn impacts the overall efficiency of housing allocation. These fluctuations highlight the challenges in maintaining an active and accurate waiting list, emphasizing the need for robust management systems to handle the complexities of application statuses.

@@

# Conclusion
This study has provided critical insights into the complex dynamics of household composition within Toronto's social housing applications from Q1 2019 to Q4 2023. By closely analyzing the data trends, the research highlights the significant challenges that families, particularly those with dependents, face in accessing stable and affordable housing. The findings underscore a marked inconsistency between the demand for social housing and the availability of accommodations, reflecting a substantial gap that persists despite varying applications rates. This persistent discrepancy indicates that current housing policies and resources may not adequately address the needs of the most vulnerable populations. As such, the study calls for a deeper and more nuanced understanding of these trends to facilitate the development of targeted policy interventions aimed at promoting housing equity and stability for families in Toronto. These efforts are crucial for not only improving living conditions but also for ensuring that housing policies evolve in alignment with the changing demographics and economic realities of the community.

Moving forward, the paper suggests several avenues for further research and policy development. It emphasizes the importance of enhancing data collection processes to capture more granular, demographic-specific information that can inform more effective housing strategies. Additionally, there is a strong recommendation for integrating adaptive policies that can swiftly respond to the fluctuations in housing demand and the specific circumstances of applicants. Such policies should focus on reducing waiting times, optimizing resource allocation, and improving the overall responsiveness of housing services. By addressing these critical issues, Toronto can make significant strides towards reducing housing inequities and ensuring that all residents have access to adequate and affordable housing. This comprehensive approach will not only address the immediate needs but also contribute to the long-term sustainability and resilience of the city’s housing system.
